---
title: "Practical"
author: "Maarten van Kessel & Ger Inberg"
date: "`r Sys.Date()`"
output: pdf_document
---

# Prerequisits
</br>
It is assumed R, RTools, Rstudio and Java are installed accourding to the [Setting up the R environment](https://ohdsi.github.io/Hades/rSetup.html) instructions.

The following packages should be installed:
1. [checkmate](https://mllg.github.io/checkmate/)
2. [DatabaseConnector](http://ohdsi.github.io/DatabaseConnector/)
3. [remotes](https://remotes.r-lib.org/)
4. [Eunomia](https://ohdsi.github.io/Eunomia/)
5. [testthat](https://testthat.r-lib.org/)
6. [knitr](https://yihui.org/knitr/)
7. [rmarkdown](https://rmarkdown.rstudio.com/)
8. [usethis](https://usethis.r-lib.org/)
9. [roxygen2](https://roxygen2.r-lib.org/articles/rd.html)

# Assignments
## 1. Create an R-project
Create a new R Package project in Rstudio (File -> New Project...). Name the project <lastname>Workshop. I.e. **vankesselWorkshop** or **inbergWorkshop**

## 2. Transform the following snippet into a function called **countPersons**
The function should take the following parameters: *connectionDetails* and *cdmSchema*
```{r, eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)
sql <- "SELECT COUNT(*) AS person_count
FROM @cdm.person;"

renderTranslateQuerySql(connection, sql, cdm = "main")
```

You can use the following function definition as a start.
```{r, eval=FALSE}
countPersons <- function(connectionDetails, cdmSchema) {
  # Your implementation
  result <- ...
  return(result)
}
```

## 3. Add function documentation to **countPersons**.
You can use the `roxygen2` package.

1. Add a function title.
2. Add a function description.
3. Add a parameter description for *connectionDetails* and *cdmSchema*.
4. Add the function to the exported functions.
5. Add a return description to what the function returns.
6. Add a **working** example.

## 4. Add parameter checking for parameters *connectionDetails* and *cdmSchema* to **countPersons**.
You can use the following functions: `checkmate::makeAssertCollection()`, `checkmate::assert_class()`, `checkmate::assert_character()`, ` checkmate::reportAssertions()`; of the `checkmate` package.

## 5. Add the used dependencies to the DESCRIPTION file.
You an use the `usethis::use_package()` function, of the `usethis` package.

## 6. Add unit testing using.
You can setup the unit testing suite with `usethis::use_testthat()`, of the `usethis` package.

You can use functions from `testthat` to test the functionality.

## 7. Add package documentation vignettes.
You can use `usethis::use_vignette()` to setup the vignette.

In the vignette showcase how **countPersons** works.


## Optional
### C. Transform the following snippet into a function called **countDrug**.
Repeat steps 3-7 for this function.
```{r, eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)
sql <- "SELECT COUNT(DISTINCT(person_id)) AS person_count
FROM @cdm.drug_exposure
INNER JOIN @cdm.concept_ancestor
  ON drug_concept_id = descendant_concept_id
INNER JOIN @cdm.concept ingredient
  ON ancestor_concept_id = ingredient.concept_id
WHERE LOWER(ingredient.concept_name) = 'celecoxib'
  AND ingredient.concept_class_id = 'Ingredient'
  AND ingredient.standard_concept = 'S';"

renderTranslateQuerySql(connection, sql, cdm = "main")
```

```{r, eval=FALSE}
countDrug <- function(connectionDetails, cdmSchema, drugName) {
  # Your implementation
  result <- ...
  return(result)
}
```


### B. Transform the following snippet into a function called **countDrugCondition**
Repeat steps 3-7 for this function.
```{r, eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)

sql <- "SELECT COUNT(DISTINCT(person_id)) AS person_count
FROM @cdm.drug_era
INNER JOIN @cdm.concept ingredient
  ON drug_concept_id = ingredient.concept_id
WHERE LOWER(ingredient.concept_name) = 'celecoxib'
  AND ingredient.concept_class_id = 'Ingredient'
  AND ingredient.standard_concept = 'S';"

renderTranslateQuerySql(connection, sql, cdm = "main")
```

```{r}
countDrugCondition <- function(connectionDetails, cdmSchema, drugName, conditionId) {
  # Your implementation
  result <- ...
  return(result)
}
```

