---
title: "Practical"
author: "Maarten van Kessel & Ger Inberg"
date: "`r Sys.Date()`"
output: pdf_document
---

# Prerequisits
It is assumed R, RTools, Rstudio and Java are installed accourding to the [*Setting up the R environment*](https://ohdsi.github.io/Hades/rSetup.html) instructions.

The following packages should be installed:

1. [*checkmate*](https://mllg.github.io/checkmate/)
2. [*knitr*](https://yihui.org/knitr/)
3. [*rmarkdown*](https://rmarkdown.rstudio.com/)
4. [*Hades*](https://ohdsi.github.io/Hades/)
    a. [*Eunomia*](https://ohdsi.github.io/Eunomia/)
    b. [*DatabaseConnector*](http://ohdsi.github.io/DatabaseConnector/)
5. [*devtools*](https://devtools.r-lib.org/)
    a. [*testthat*](https://testthat.r-lib.org/)
    b. [*roxygen2*](https://roxygen2.r-lib.org/)
    c. [*remotes*](https://remotes.r-lib.org/)
    d. [*usethis*](https://usethis.r-lib.org/)
  
```{r Installation, eval=FALSE}
# Install packages from CRAN
install.packages(c("devtools", "checkmate", "knitr", "rmarkdown", "DatabaseConnector"))

# Install packages from OHDSI
remotes::install_github("OHDSI/Eunomia")
```
  
# Assignments
## 1. Create an R-project
1. Create a new R Package project in Rstudio (File -> New Project...). Name the project <firstname initial><Lastname>Workshop. I.e. **mVankesselWorkshop** or **gInbergWorkshop**.
2. Remove the **hello.R** and **man/hello.Rd** files.
3. Create a new R-file named **getCounts.R**.

## 2. Transform the following snippet into a function called **countPersons**
The function should take the following parameters: *connectionDetails* and *cdmSchema*
```{r snippet_2, eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)
sql <- "SELECT COUNT(*) AS person_count
FROM @cdm.person;"

renderTranslateQuerySql(connection, sql, cdm = "main")
```

You can use the following function definition as a start.
```{r constructor_2, eval=FALSE}
countPersons <- function(connectionDetails, cdmSchema) {
  # Your implementation
  result <- ...
  return(result)
}
```

Make sure you close the database connection regardless if the function throws an error or not.

## 3. Add function documentation to **countPersons**.
You can use the `roxygen2` package.

1. Add a function title.
2. Add a function description.
3. Add a parameter description for *connectionDetails* and *cdmSchema*.
4. Import `DatabaseConnector` and `checkmate` in your package.
5. Add the function to the exported functions.
6. Add a return description to what the function returns.
7. Add a **working** example.
8. Use `roxygen2::roxygenise()` to add the documentation files.

## 4. Add parameter checking for parameters *connectionDetails* and *cdmSchema* to **countPersons**.
You can use various functions of `checkmate`.

## 5. The DESCRIPTION file.

1. Add the used dependencies
2. Add a title in *Title Case*.
3. Add an author.
4. Add an maintainer.
5. Add an encoding.

You can use various functions of `usethis`.

## 6. Add unit testing.
You can setup the unit testing suite with `usethis::use_testthat()`, of the `usethis` package.

You can use functions from `testthat` to test the functionality.

## 7. Add package documentation vignettes.
You can use `usethis::use_vignette()` to setup the vignette.

In the vignette showcase how **countPersons** works, and what the results mean.

## 8. R-Check
To verify that the package is installable and use able use `devtools::check()`. This may take 1-5 minutes. Investigate, and resolve any *Error*, *Warning*, and *NOTE* messages that you encounter.

## Optional
### A. Code review
Review and compare your implementation with the  [*answers*](https://github.com/ginberg/erasmusmc-workshop/tree/answers).

### B. Transform the following snippet into a function called **countDrug**.
Repeat steps 3-8 for this function.

Use the following snippet to create a function that counts the number of people that are taking a drug by name.
```{r snippet_B, eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)
sql <- "SELECT COUNT(DISTINCT(person_id)) AS person_count
FROM @cdm.drug_exposure
INNER JOIN @cdm.concept_ancestor
  ON drug_concept_id = descendant_concept_id
INNER JOIN @cdm.concept ingredient
  ON ancestor_concept_id = ingredient.concept_id
WHERE LOWER(ingredient.concept_name) = 'celecoxib'
  AND ingredient.concept_class_id = 'Ingredient'
  AND ingredient.standard_concept = 'S';"

renderTranslateQuerySql(connection, sql, cdm = "main")
```

```{r constructor_B, eval=FALSE}
countDrug <- function(connectionDetails, cdmSchema, drugName) {
  # Your implementation
  result <- ...
  return(result)
}
```


### C. Transform the following snippet into a function called **countDrugCondition**
Repeat steps 3-7 for this function.

Use the following snippet to count the number of people, with a condition (by concept ID), taking a drug (by name).
```{r snippet_C, eval=FALSE}
library(DatabaseConnector)

connection <- connect(connectionDetails)

sql <- "SELECT COUNT(*) AS diagnose_count
FROM @cdm.drug_era
INNER JOIN @cdm.concept ingredient
  ON drug_concept_id = ingredient.concept_id
INNER JOIN @cdm.condition_occurrence
  ON condition_start_date >= drug_era_start_date
    AND condition_start_date <= drug_era_end_date
INNER JOIN @cdm.concept_ancestor
  ON condition_concept_id = descendant_concept_id
WHERE LOWER(ingredient.concept_name) = 'celecoxib'
  AND ingredient.concept_class_id = 'Ingredient'
  AND ingredient.standard_concept = 'S'
  AND ancestor_concept_id = 192671;"

renderTranslateQuerySql(connection, sql, cdm = "main")
```

```{r constructor_C, eval=FALSE}
countDrugCondition <- function(connectionDetails, cdmSchema, drugName, conditionId) {
  # Your implementation
  result <- ...
  return(result)
}
```
